Who introduced this bug? \\ It may not have been caused by the previous commit!

### 2016-01-29 paperMSR.tex
Retocando las tablas y alguna cosa más. Luego espero seguir

\title{Who introduced this bug? \\ It may not have been in the previous commit!}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

The remainder of this paper is structured as follows. Next, we present the current body of knowledge in section~\ref{sec:related}. Section~\ref{sec:methodology} describes the methodology used to identify the moment in which the bug was introduced in the source code, followed by the results obtained after applying our approach to a selection of OpenStack bug fixes in Section~\ref{sec:results}. Section~\ref{sec:discussion} answers the research questions and discusses potential applications and improvements of our approach. After reporting the limitations and threats to validity in Section~\ref{sec:threats}, we draw some conclusions in Section~\ref{sec:conclusions}.

@@

All data needed to analyze when the bug was introduced can be obtained from the issue tracking systems and the code review systems used generally by free/open source software (FOSS) projects. In our analysis, we have used OpenStack, a cloud computing platform with a huge developing community (more than 5,000 developers) and significant industrial support from several major companies such as Red Hat, Intel, IBM, HP, etc. OpenStack, as many other FOSS projects nowadays, uses Launchpad\footnote{\url{https://launchpad.net/openstack}} as issue tracking system, and Gerrit\footnote{\url{https://review.openstack.org/}} to support their code review process.

@@

@@

The next part is focusing on analyzing the previous commit in the \textit{Bug Report} group. For that, we had to analyze the lines involved in the bug fix, in the commit parent of the bug fix commit, and be sure that the lines was inserted/modified in the previous commit. This, way we can sure that the previous commit didn't copied any line that contained the bug, because in this case, the previous commit is not responsible to cause the bug.

The analysis was done manually, and we used \textit{git blame} to see all the previous commit in each line of a involved file. Also, we used \textit{diff} to see all the differences between two files, in our case, the file is going to be the same but in different moment inside the control version system used.

Te procedure following in each file involved in a fix bug is describing below;
\begin{enumerate}
  \item git checkout \textit{commit that fix the bug}, git blame \textit{file involved}. In this step we can see the lines added, modified or deleted by the commit that fix the bug.
  \item git checkout \textit{parent of commit that fix the bug}, git blame \textit{file involved}. In this step we can see all the previous commits involved in the different lines touched in the fix bug.
  \item git checkout \textit{parent of previous commit}, git blame \textit{file involved}. In this step we can ensure that the previous commit inserted these lines.
\end{enumerate}

Finally we need to discard some noise presents in our final results according to the responsibility of the previous commit inserting the bug in the code source. Due to they were not responsible for cause the bug, we delete those previous commit which presents the following criteria;
\begin{itemize}
    \item Blank lines
    \item Format changes
    \item Copied lines
    \item Changes in the comment.
    \item Updates in the version of a file. 
\end{itemize}

@@

We validate our methodology analyzing 459 tickets in OpenStack. OpenStack was particularly of interest because of its continuously evolving due to its very active community. Although its short life, only five years, more than 5 thousand of researchers and more than 233 thousand of commits with more than 2 Million of lines of code have contributed in the development of the project\footnote{\url{http://activity.openstack.org/dash/browser/}}. Furthermore, all history is saved and available in a version control system, being able to access to its issue tracking system\footnote{\url{https://launchpad.net/openstack}} and the source code review\footnote{\url{https://review.openstack.org/}}

OpenStack is composed by 9 projects, but we only focused in the four of them, Nova, Cinder, Neutron and Horizon as we can see in table \ref{tab:OpenStack} these projects are really actives during all their history and in the last year.
\begin{table}[htb]
\centering
%\begin{center} {\footnotesize
\begin{tabular}{lrr}
\toprule[0.3mm]%{\smallskip}
  & All History  & Last Year (2015) \\\hline
Nova    \kern 1pc & 14,558 & 3,283 \\
Fuel    \kern 1pc & 9,139 & 5,123 \\
Netron  \kern 1pc & 8,452 & 3,855 \\
Horizon \kern 1pc & 4,871 & 1,994 \\
Cinder  \kern 1pc & 4,556 & 1,832 \\
Keystone\kern 1pc & 4,874 & 1,795  \\
Heat    \kern 1pc & 6,395 & 2,372  \\
Glance  \kern 1pc & 2,651 & 723 \\
Tempest \kern 1pc & 4,141 & 1,312 \\
\bottomrule[0.3mm]
\end{tabular} %}
\caption{ Commits per Project in OpenStack}
\label{tab:OpenStack}

@@

\begin{table}
\centering
%\begin{center} {\footnotesize
\begin{tabular}{lllll}
\toprule[0.3mm]%{\smallskip}
  & Bug Report & Not Bug Report & Undecided & Total \\\hline
R1  & (184)55\% & (115)34\% & (35)11\% & 334 \\
R2  & (188)76\% & (54)22 \% & (7)3\% & 249 \\
R3 & (188)56\% & (116)35\% & (30)9\% & 334 \\
Finally & (209)72\% & (74)25\% & (9)3\% & 292 \\
\bottomrule[0.3mm]
\end{tabular} %}
\caption{ Statistics of each researcher in the classification}
\label{tab:2}
%\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center} {\footnotesize
\begin{tabular}{llllll}
\toprule[0.3mm]%{\smallskip}
  & Nova & Cinder & Horizon & Neutron & Total\\\hline
R1\&R2   & (44)70\% & (40)77\%  & (37)60\% & -        & 68\% \\
R1\&R3   &  -        & (46)73\%  & (48)76\% & (26)62\% & 71 \% \\
R2\&R3   & (41)66\% & (10)100\% & -         & -        &  71\% \\
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Concordance between each developer in each repository}
\label{tab:3}
\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center} {\footnotesize
\begin{tabular}{lcc}
\toprule[0.3mm]
  & \multicolumn{1}{c}{Before} & \multicolumn{1}{c}{After} \\
  & \multicolumn{1}{c}{Deleting Noise} & \multicolumn{1}{c}{Deleting Noise} \\\hline
\raisebox{1ex}{Responsible} & 152 & 152 \\[0ex]
\raisebox{1ex}{Not responsible} & 154 & 114 \\[0ex]
\raisebox{1ex}{Undecided} & 42 & 42 \\[0ex]
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Responsibility of each previous commit before and after deleting the noise in the results}
\label{tab:responsability}
\end{center}
\end{table}

>>>>

\title{Who introduced this bug? \\ It may not have been caused by the previous commit!}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

The remainder of this paper is structured as follows. Next, we present the current body of knowledge in section~\ref{sec:related}. Section~\ref{sec:methodology} describes the methodology used to identify the moment in which the bug was introduced in the source code, followed by the results obtained after applying our approach to a selection of OpenStack bug fixes in Section~\ref{sec:results}. Section~\ref{sec:discussion} answers the research questions and discusses potential applications and improvements of our approach. After reporting the limitations and threats to validity in Section~\ref{sec:threats}, we draw some conclusions and point out some potential future work in Section~\ref{sec:conclusions}.

@@

All data needed to analyze when the bug was introduced can be obtained from the issue tracking systems and the code review systems used generally by free/open source software (FOSS) projects. In our analysis, we have focused on Launchpad\footnote{\url{https://launchpad.net/}} as issue tracking system, and Gerrit\footnote{\url{https://www.gerritcodereview.com/}} as code review supporting tool, as they are widely used by FOSS projects nowadays, but our methodology should be generalizable to any such tool.

@@

FIXME: aqui podemos poner un pantallazo de la interfaz web.

@@

The next part is focused on analyzing the previous commit exclusively in the \textit{Bug Report} group. For that, we had to analyze the lines involved in the bug fix, in the \emph{parent} commit of the bug fix commit, and be sure that the lines were inserted/modified in the previous commit. This way we can be sure that the previous commit did not copy lines that contained the bug, because in that case, the previous commit is not responsible to cause the bug. FIXME: aclarar este parrafo!

The analysis was done manually. We used \textit{git blame} to see the previous commit for each line of the involved file. Also, we used \textit{diff} to see the differences between the two files, in our case as the file is going to be the same, between the file in two different moments in the control version system.

The procedure for each file involved in a bug fix is as follows:

\begin{enumerate}
  \item git checkout \textit{commit that fixed the bug}, git blame \textit{file involved}. In this step we can see the lines added, modified or deleted by the commit that fixed the bug.
  \item git checkout \textit{parent of commit that fix the bug}, git blame \textit{file involved}. In this step we can see the previous commits for the different lines touched in the fixed bug.
  \item git checkout \textit{parent of previous commit}, git blame \textit{file involved}. With this step we can ensure that the previous commit inserted these lines.
\end{enumerate}

Finally we need to discard some noise presents in our final results according to the responsibility of the previous commit inserting the bug in the code source. Due to they were not responsible for cause the bug, we delete those previous commit which presents the following criteria:

\begin{itemize}
    \item Blank lines
    \item Format changes
    \item Copied lines
    \item Changes in the comment.
    \item Updates in the version of a file. 
\end{itemize}

@@

We validate our methodology analyzing 459 tickets from OpenStack. OpenStack is a cloud computing platform with a huge developing community (more than 5,000 developers) and significant industrial support from several major companies such as Red Hat, Intel, IBM, HP, etc. OpenStack was particularly of interest because of its continuously evolving due to its very active community. Currently it has more than 233,000 commits with more than 2 million lines of code \footnote{\url{http://activity.openstack.org/dash/browser/}}. All its history is saved and available in a version control system, as well as its issue tracking system (Launchpad\footnote{\url{https://launchpad.net/openstack}}) and the source code review system (Gerrit\footnote{\url{https://review.openstack.org/}}).

OpenStack is composed by 9 projects, but we only focused in four of them: Nova, Cinder, Neutron and Horizon. As can be seen from Table~\ref{tab:OpenStack}, these projects have been very active during their entier history, and in the last year.

\begin{table}[htb]
\centering
%\begin{center} {\footnotesize
\begin{tabular}{lrr}
\toprule[0.3mm]%{\smallskip}
  & All History  & Last Year (2015) \\\hline
Nova    \kern 1pc & 14,558 & 3,283 \\
Fuel    \kern 1pc & 9,139 & 5,123 \\
Netron  \kern 1pc & 8,452 & 3,855 \\
Horizon \kern 1pc & 4,871 & 1,994 \\
Cinder  \kern 1pc & 4,556 & 1,832 \\
Keystone\kern 1pc & 4,874 & 1,795  \\
Heat    \kern 1pc & 6,395 & 2,372  \\
Glance  \kern 1pc & 2,651 & 723 \\
Tempest \kern 1pc & 4,141 & 1,312 \\
\bottomrule[0.3mm]
\end{tabular} %}
\caption{ Commits per Project in OpenStack}
\label{tab:OpenStack}

@@

\begin{table}
\centering
%\begin{center} {\footnotesize
\begin{tabular}{lrrrr}
\toprule[0.3mm]%{\smallskip}
  & Bug Report & Not Bug Report & Undecided & Total \\\hline
R1  & (184) 55\% & (115)34\% & (35) 11\% & 334 \\
R2  & (188) 76\% & (54)22 \% & (7) 3\% & 249 \\
R3 & (188) 56\% & (116) 35\% & (30) 9\% & 334 \\
Finally & (209) 72\% & (74) 25\% & (9) 3\% & 292 \\
\bottomrule[0.3mm]
\end{tabular} %}
\caption{ Statistics of each researcher in the classification}
\label{tab:2}
%\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center} {\footnotesize
\begin{tabular}{lrrrrr}
\toprule[0.3mm]%{\smallskip}
  & Nova & Cinder & Horizon & Neutron & Total\\\hline
R1\&R2   & (44) 70\% & (40) 77\%  & (37) 60\% & -        & 68\% \\
R1\&R3   &  -        & (46) 73\%  & (48) 76\% & (26) 62\% & 71\% \\
R2\&R3   & (41) 66\% & (10) 100\% & -         & -        &  71\% \\
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Concordance between each developer in each repository}
\label{tab:3}
\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center} {\footnotesize
\begin{tabular}{lrr}
\toprule[0.3mm]
  & \multicolumn{1}{c}{Before} & \multicolumn{1}{c}{After} \\
  & \multicolumn{1}{c}{Deleting Noise} & \multicolumn{1}{c}{Deleting Noise} \\\hline
\raisebox{1ex}{Responsible} & 152 & 152 \\[0ex]
\raisebox{1ex}{Not responsible} & 154 & 114 \\[0ex]
\raisebox{1ex}{Undecided} & 42 & 42 \\[0ex]
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Responsibility of each previous commit before and after deleting the noise in the results}
\label{tab:responsability}
\end{center}
\end{table}
